---
title: "Reproducible Research: Peer Assessment 1"
author: "Adam Weissman"
date: "1 September 2018"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,cache = TRUE)
```

## Dependencies
The following libraries are used by this script:
```{r libraries,echo=TRUE}
library(knitr)
library(ggplot2)
library(dplyr)
library(readr)
library(lubridate)

```

## Constants
The following contstants will be  are used by this script:
```{r constants,echo=TRUE}
    dlzip="activity.zip"
    dlurl="https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
    datafilename="activity.csv"

```




## Loading and preprocessing the data
Download the file if it doesn't already exist in the workspace
```{r download,echo=TRUE}
if(!file.exists(dlzip)){
    download.file(dlurl,dlzip)
    unzip(dlzip)
}
```
Read data into R and format the date variable data
```{r prep, echo=TRUE,results='hide'}
    activity<-read_csv(datafilename)
    activity$date<-ymd(activity$date)
```


## What is mean total number of steps taken per day?
```{r mean,echo=FALSE}
    actAgg<-aggregate(steps~date,data=activity,FUN=sum,na.action=na.omit)
    g<-ggplot(data = actAgg,aes(x=steps))
    g+geom_histogram(bins=30)+geom_vline(xintercept=mean(actAgg$steps), color="red")+geom_vline(xintercept=median(actAgg$steps),color="blue")+labs(x="Steps taken per Day",y="Count",title="Count of Steps taken per Day")
    
    print(paste("Mean  : ", floor(mean(actAgg$steps))))
    print(paste("Median: ", median(actAgg$steps)))
    
```

## What is the average daily activity pattern?
```{r daily,echo=FALSE}
    actDaily<-activity%>%group_by(interval)%>%summarise(mStep=mean(steps,na.rm=TRUE))
    g<-ggplot(data = actDaily,aes(x=interval,y=mStep))
    g+geom_line()+labs(title="Average Daily Activity",x="Time of day interval",y="Average Steps at Time of Day")+geom_vline(xintercept=actDaily[which(actDaily$mStep==max(actDaily$mStep)),]$interval,color="red")
    
    print(paste("Most Active Average Time Interval  : ", actDaily[which(actDaily$mStep==max(actDaily$mStep)),]$interval))
    
```


## Imputing missing values
Take the average step count for the time interval bin and assign it to the missing measurements
```{r nas, echo=FALSE}
    print(paste("Number of missing values: ", sum(is.na(activity$steps))))
    actImp<-activity    
    actImp[is.na(activity$steps),]$steps<-actDaily[match(activity[is.na(activity$steps),]$interval,actDaily$interval),]$mStep

    actImpAgg<-aggregate(steps~date,data=actImp,FUN=sum,na.action=na.omit)
    g<-ggplot(data = actImpAgg,aes(x=steps))

    g+geom_histogram(bins=30)+geom_vline(xintercept=mean(actImpAgg$steps), color="red")+geom_vline(xintercept =median(actImpAgg$steps),color="blue")+labs(x="Number of Steps",y="Count of Days with Step totals",title="Count of Steps taken per Day") 
    print(paste("Mean  : ", floor(mean(actImpAgg$steps))))
    print(paste("Median: ", floor(median(actImpAgg$steps))))

```

## Are there differences in activity patterns between weekdays and weekends?
```{r weekend,echo=FALSE}
    actImp<-actImp%>%mutate(dow=factor(wday(actImp$date),levels=1:7,c("Weekend",rep("Weekday",5),"Weekend")))
    actImpDaily<-actImp%>%group_by(dow,interval)%>%summarise(mStep=mean(steps))
    g<-ggplot(actImpDaily,aes(x=interval,y=mStep))
    g+geom_line()+facet_grid(dow~.)+labs(x="Time Interval",y="Average Steps",title="Average Daily Activity comparing Weekend and Weekday Activity")
    print(paste("Mean steps weekday:",floor(sum(actImpDaily[actImpDaily$dow=="Weekday",]$mStep))))
    print(paste("Mean steps weekend:",floor(sum(actImpDaily[actImpDaily$dow=="Weekend",]$mStep))))
```


